---
title: "Sunshine"
author: "Subam"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This R Markdown document explores sunshine duration data from various countries and cities. The goal is to uncover insights and patterns related to sunshine patterns and their potential implications.

## Data Exploration

### Loading the Data
```{r}
library(tidyverse)
sunshine_data <- read.csv("E:\\Portfolio_projects\\sunshine.csv")
# Print the number of rows and columns
dim(sunshine_data)
```
### Check for missing values and duplicated values
```{r}
# Checking for missing values
sum(is.na(sunshine_data))  # Count total missing values
colSums(is.na(sunshine_data))  # Count missing values per column
# Checking for duplicates
sum(duplicated(sunshine_data))  # Count duplicate rows
```
### Unique Countries and Cities
```{r}
# Find the number of unique countries
n_countries <- length(unique(sunshine_data$Country))

# Find the number of unique cities
n_cities <- length(unique(sunshine_data$City))

# Print the results
cat("Number of unique countries:", n_countries, "\n")
cat("Number of unique cities:", n_cities)
```
### Total Sunshine for a year in different Country or City
```{r}
# Find the minimum and maximum years
min_year <- min(sunshine_data$Year)
max_year <- max(sunshine_data$Year)

# Print the range of years
cat("Total sunshine for a year per country:", min_year, "-", max_year)
```
### Average Sunshine Duration per month
```{r}
# Calculate monthly averages across all countries/cities
monthly_averages <- colMeans(sunshine_data[, 3:14])  # Exclude first two columns (Country and City)

# Print the results
month_names <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
cat("Average Sunshine Duration by Month:", "\n")
sapply(month_names, FUN = function(month) paste0(month, ": ", round(monthly_averages[match(month, month_names)], 2), " hours"))
```
# sapply() loops through month names and prints the corresponding average sunshine duration rounded to two decimal places.

###  Country/City with Highest Average Sunshine
```{r}
# Calculate mean sunshine duration for each country/city
country_city_averages <- rowMeans(sunshine_data[, 3:14])

# Find the maximum value and its corresponding row index
max_index <- which.max(country_city_averages)

# Extract the country and city from the row index
max_country <- sunshine_data$Country[max_index]
max_city <- sunshine_data$City[max_index]

# Print the results
cat("\nCountry/City with Highest Average Sunshine:", "\n")
cat(paste0(max_country, ", ", max_city, ": ", round(country_city_averages[max_index], 2), " hours"))
```
### Which country has the most significant difference in sunshine duration between summer and winter?
```{r}
sunshine_data <- read.csv("E:\\Portfolio_projects\\sunshine.csv")
# Define the months corresponding to summer and winter
summer_months <- c("Jun", "Jul", "Aug")
winter_months <- c("Dec", "Jan", "Feb")

# Calculate the average sunshine duration for summer and winter
sunshine_data <- sunshine_data %>%
  group_by(Country) %>%
  summarize(
    SummerDuration = mean(c_across(all_of(summer_months)), na.rm = TRUE),
    WinterDuration = mean(c_across(all_of(winter_months)), na.rm = TRUE),
    Difference = SummerDuration - WinterDuration
  ) %>%
  arrange(desc(Difference))  # Arrange in descending order of difference

# Display the top country with the most significant difference
top_country <- sunshine_data[1, "Country"]
print(paste("The country with the most significant difference is:", top_country))
```
### Months with Highest and Lowest Sunshine
```{r}
# Identify the month indices with maximum and minimum averages
highest_month_index <- which.max(monthly_averages)
lowest_month_index <- which.min(monthly_averages)

# Extract the corresponding month names
highest_month <- month_names[highest_month_index]
lowest_month <- month_names[lowest_month_index]

# Print the results
cat("\nMonth with Highest Sunshine:", highest_month, "\n")
cat("Month with Lowest Sunshine:", lowest_month)
```

### Which cities have the most consistent sunshine duration throughout the year (i.e., least variation between months)?
```{r}
sunshine_data <- read.csv("E:\\Portfolio_projects\\sunshine.csv")
# Calculate the coefficient of variation (CV) for each city
sunshine_variation <- sunshine_data %>%
  group_by(Country, City) %>%
  summarize(
    CV = sd(c_across(starts_with("Jan", ignore.case = TRUE):starts_with("Dec", ignore.case = TRUE)), na.rm = TRUE) /
         mean(c_across(starts_with("Jan", ignore.case = TRUE):starts_with("Dec", ignore.case = TRUE)), na.rm = TRUE)
  ) %>%
  arrange(CV)  # Arrange in ascending order of CV

# Display cities with the least variation
top_cities <- head(sunshine_variation, 5)  # Display the top 5 cities
print(top_cities)
```
# These are the top 5 cities which experience the most sunshine in a given year.

### Which cities have the most variable sunshine duration throughout the year (i.e., greatest variation between months)?
```{r}
# Calculate the coefficient of variation (CV) for each city
sunshine_variation <- sunshine_data %>%
  group_by(Country, City) %>%
  summarize(
    CV = sd(c_across(starts_with("Jan", ignore.case = TRUE):starts_with("Dec", ignore.case = TRUE)), na.rm = TRUE) /
         mean(c_across(starts_with("Jan", ignore.case = TRUE):starts_with("Dec", ignore.case = TRUE)), na.rm = TRUE)
  ) %>%
  arrange(desc(CV))  # Arrange in descending order of CV

# Display cities with the most variation
top_cities <- head(sunshine_variation, 5)  # Display the top 5 cities
print(top_cities)
```
# These are the top 5 cities that experience least amount of sunshine in a year.

### Interactive Map
```{r}
# Load necessary libraries
library(tidyverse)
library(leaflet)

# Reshape the data for plotting
melted_data <- pivot_longer(sunshine_data, cols = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"), names_to = "Month", values_to = "SunshineDuration")

# Convert 'Month' column to a factor with proper ordering
melted_data$Month <- factor(melted_data$Month, levels = month.name)

# Read geolocation data from CSV file
geolocation_data <- read.csv("E:\\Portfolio_projects\\latandlon.csv") 

# Identify common columns for merging
common_columns <- intersect(names(geolocation_data), names(melted_data))

# Check if the common columns are present
if (length(common_columns) == 0) {
  stop("No common columns found for merging. Check column names in both datasets.")
}

# Merge datasets based on common columns
merged_data <- merge(melted_data, geolocation_data, by = common_columns, all.x = TRUE)

# Calculate average sunshine duration for each location
avg_sunshine <- merged_data %>%
  group_by(Country, City, Latitude, Longitude) %>%
  summarise(AvgSunshineDuration = mean(SunshineDuration, na.rm = TRUE))

# Filter out rows with missing or invalid lat/lon values
avg_sunshine <- avg_sunshine[complete.cases(avg_sunshine$Latitude, avg_sunshine$Longitude), ]

# Create an interactive map using leaflet without marker clustering
interactive_map <- leaflet(avg_sunshine) %>%
  addTiles() %>%
  addCircleMarkers(
    lat = ~Latitude,
    lng = ~Longitude,
    radius = ~AvgSunshineDuration/50,  # Adjust the radius based on your data
    color = "orange",
    stroke = FALSE,
    fillOpacity = 0.7,
    popup = ~paste(Country, City, "<br>", "Avg. Sunshine Duration:", round(AvgSunshineDuration, 2), "hours")
  ) %>%
  addLegend(
    "bottomright",
    colors = "orange",
    labels = "Avg. Sunshine Duration",
    opacity = 1
  )

# Display the interactive map
interactive_map
```

